Adolescent well-being in the digital age

### 2022-02-01 clean.Rmd
Take out gwpi

```{r gallup-codebook}
#| echo: false
#| eval: false
# Understand variables & values
# Codes NOTE 1: yes 2: no!
read_spss(
  "data-raw/Gallup/The_Gallup_101521.sav",
  n_max = 1,
  col_select = c(
    WPID, WP1219,
    WP16,
    WP60, WP61, WP63, WP65, WP67,
    WP68, WP69, WP70, WP71, WP74,
    WP16056
  )
) %>%
  generate_dictionary()
```

@@

```{r gallup-clean}
# First liberate the data from the slow SPSS file and save to disk
gwp_path <- "data-raw/Gallup/gwp-processed.rds"
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    "data-raw/Gallup/The_Gallup_101521.sav",
    col_select = c(
      YEAR_CALENDAR,
      COUNTRYNEW,
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74,
      WP16056
    )
  )

@@

gwp <- gwp %>%
  clean_names() %>%
  transmute(
    country = countrynew,
    year = year_calendar,
    id = wpid,
    sex = factor(wp1219, levels = c(2, 1), labels = c("Female", "Male")),
    age = wp1220,
    # Don't know, refused, and missing values
    Life_satisfaction = if_else(between(wp16, 0, 10), wp16, NaN),
    across(wp60:wp74, ~ if_else(between(., 1, 2), ., NaN)),
    # Also reverse the weird 1: yes 2: no coding here
    across(wp60:wp74, ~ 3 - .),
    # Internet (fix NAs and reverse 1:yes 2:no coding to 1:no 2:yes)
    gwpi = 2 - if_else(between(wp16056, 1, 2), wp16056, NaN)
  )

@@

```{r gallup-summarise}
gwp <- gwp %>%
  # Aggregate internet
  group_by(country, year, sex, age) %>% 
  mutate(gwpi = mean(gwpi, na.rm = TRUE)) %>% 
  pivot_longer(
    c(
      Life_satisfaction,
      Negative_experiences,
      Positive_experiences
    ),
    names_to = "outcome", values_to = "val"
  ) %>%
  group_by(country, year, sex, age, outcome, gwpi) %>%
  summarise(
    n = n(),
    se = sd(val, na.rm = TRUE) / sqrt(n),
    val = mean(val, na.rm = TRUE)
  ) %>%
  ungroup()
```

@@

```{r}
# Create if doesn't exist
gwp_mock_path <- "data-raw/gwp-MOCK.rds"
if (!file.exists(gwp_mock_path)) {
  set.seed(101010)
  distinct(gwp, outcome, country, year, sex, age, gwpi) %>% 
    mutate(
      se = rnorm(
        nrow(.), 
        mean(gwp$se, na.rm = TRUE), 
        sd(gwp$se, na.rm = TRUE)
      ),
      val = rnorm(
        nrow(.), 
        mean(gwp$val, na.rm = TRUE), 
        sd(gwp$val, na.rm = TRUE)
      )
    ) %>% 
    write_rds(gwp_mock_path, compress = "xz")
}
# Load if actual data doesn't exist
if (!dir.exists("data-raw/Gallup")) {
  cat("Using synthetic GWP data.")
  gwp <- read_rds(gwp_mock_path)
} else {
  cat("Using actual GWP data.")
}
```

@@

# Reshape to easily impute all variables
itu_long <- itu %>%
  pivot_longer(c(internet, fixed, mobile))

@@

## GWP summary table

```{r gwp-summary-table}
#| column: page-right
#| tbl-cap: "Mean internet user percentages across regions and time from the GWP data."
#| label: tbl-gwpsummary

gwp %>%
  filter(year >= 2015) %>% 
  group_by(region, country, year) %>%
  summarise(gwpi = mean(gwpi, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(gwpi = gwpi * 100) %>%
  mutate(region = str_replace(region, "_", " ")) %>%
  arrange(year, region, country) %>% 
  pivot_wider(names_from = year, values_from = gwpi) %>%
  select(-country) %>%
  tbl_summary(
    by = region,
    missing = "no",
    statistic = list(all_continuous() ~ "{mean}% ({N_nonmiss})")
  ) %>%
  add_overall(col_label = "Total") %>%
  as_kable_extra(
    caption = "GWP Internet use summary (mean over demographics)"
  ) %>%
  row_spec(0, bold = TRUE) %>%
  kable_custom()
```

@@

```{r}
write_rds(itu, "data/itu.rds")
write_rds(gwp, "data/gwp.rds")
write_rds(gbd, "data/gbd.rds")
```

>>>>

```{r gallup-codebook}
#| echo: false
#| eval: false
# Understand variables & values
# Codes NOTE 1: yes 2: no!
read_spss(
  "data-raw/Gallup/The_Gallup_101521.sav",
  n_max = 1,
  col_select = c(
    WPID, WP1219,
    WP16,
    WP60, WP61, WP63, WP65, WP67,
    WP68, WP69, WP70, WP71, WP74
  )
) %>%
  generate_dictionary()
```

@@

```{r gallup-clean}
# First liberate the data from the slow SPSS file and save to disk
gwp_path <- "data-raw/Gallup/gwp-processed.rds"
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    "data-raw/Gallup/The_Gallup_101521.sav",
    col_select = c(
      YEAR_CALENDAR,
      COUNTRYNEW,
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74
    )
  )

@@

gwp <- gwp %>%
  clean_names() %>%
  transmute(
    country = countrynew,
    year = year_calendar,
    id = wpid,
    sex = factor(wp1219, levels = c(2, 1), labels = c("Female", "Male")),
    age = wp1220,
    # Don't know, refused, and missing values
    Life_satisfaction = if_else(between(wp16, 0, 10), wp16, NaN),
    across(wp60:wp74, ~ if_else(between(., 1, 2), ., NaN)),
    # Also reverse the weird 1: yes 2: no coding here
    across(wp60:wp74, ~ 3 - .)
  )

@@

```{r gallup-summarise}
gwp <- gwp %>%
  pivot_longer(
    c(
      Life_satisfaction,
      Negative_experiences,
      Positive_experiences
    ),
    names_to = "outcome", values_to = "val"
  ) %>%
  group_by(country, year, sex, age, outcome) %>%
  summarise(
    n = n(),
    se = sd(val, na.rm = TRUE) / sqrt(n),
    val = mean(val, na.rm = TRUE)
  ) %>%
  ungroup()
```

@@

```{r}
# Create if doesn't exist
gwp_mock_path <- "data-raw/gwp-MOCK.rds"
if (!file.exists(gwp_mock_path)) {
  set.seed(101010)
  # TODO
}
# Load if actual data doesn't exist
if (!dir.exists("data-raw/Gallup")) {
  cat("Using synthetic GWP data.")
  gwp <- read_rds(gwp_mock_path)
} else {
  cat("Using actual GWP data.")
}
```

@@

# Reshape to easily impute all variables
itu_long <- itu %>%
  pivot_longer(c(internet, mobile))

@@

## Prepare model data

-   Create lagged and centered predictors
-   Transform edge values of internet variable to fit beta
    -   2 values were changed (0 in Timor-Leste in 2002, and 1 in UAE 2020)
-   Join outcome and predictor tables
-   Center year on 2010 (roughly in the middle)
-   Ensure that age and sex are treated with contrast codes

```{r model-prepare-data}
# Create lagged predictors
itu <- itu %>% 
  arrange(region, country, year) %>%
  group_by(region, country) %>%
  mutate(
    across(
      c(internet, mobile),
      list(`1` = ~ lag(., 1)),
      .names = "{str_sub(.col, 1, 1)}{.fn}"
    )
  ) %>% 
  ungroup()

# Centering predictors within and between countries
itu <- itu %>%
  group_by(country) %>%
  mutate(
    i1_cb = mean(i1, na.rm = TRUE), # Country average
    i1_cw = i1 - i1_cb, # Year-specific deviation from country average
    m1_cb = mean(m1, na.rm = TRUE),
    m1_cw = m1 - m1_cb,
  ) %>%
  # Grand-mean center the country averages
  ungroup() %>% 
  mutate(
    across(c(i1_cb, m1_cb), ~ . - mean(., na.rm = T))
  )

# Transform internet edge cases
itu <- itu %>%
  mutate(internet = if_else(internet == 0, .0000001, internet)) %>%
  mutate(internet = if_else(internet == 1, .9999999, internet))

# Stack outcome data
dat <- bind_rows(gwp, gbd)

# Merge predictors to outcomes
dat <- left_join(dat, itu)

# Center and scale year
dat <- dat %>%
  mutate(year = (year - 2010) / 10)

# Identify values to use in mv models
# Indicate unique ITU values for MV model
dat <- dat %>%
  # Unique values exist for each country-year, but must be given for each
  # outcome
  group_by(region, country, year, outcome) %>%
  mutate(itu = 1:n() == 1) %>%
  ungroup()

# Ensure that categorical variables are factors
dat <- dat %>%
  mutate(across(c(sex, age), factor))

# Model outcomes on the percentage (0-100) scale. Note internet is still 0-1
dat <- dat %>%
  mutate(across(c(val, se), ~ . * 100))

# Arrange on outcome
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
dat <- dat %>%
  mutate(outcome = factor(outcome, levels = oo)) %>% 
  arrange(outcome, region, country, year)

# Selfharm models don't converge with SEs
dat <- dat %>% 
  mutate(se = if_else(outcome == "Selfharm", 0, se))
```

@@

```{r}
write_rds(dat, "data/data-all.rds")
write_rds(itu, "data/itu.rds")
write_rds(gwp, "data/gwp.rds")
write_rds(gbd, "data/gbd.rds")
```
